---
title: "153project"
author: "Jiyoon Clover Jeong and Jin Kweon"
date: "3/24/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggfortify)
library(stats)
library(TSA)
library(forecast)
library(tsoutliers)
library(astsa)
library(tseries)
library(dynlm)
library(stats)
library(randtests) #turning.point.test
library(reshape2)
library(gridExtra)
options(scipen=10000) #No scientific notation
```

#EDA
```{r EDA}
hawaii <- read.csv("../data/Carbon_Hawaii.csv")
dim(hawaii)
hawaii.ts <- ts(hawaii$Carbondioxide, frequency=12, start=c(1959, 1), end=c(1990, 12))
anyNA(hawaii.ts)

plot(hawaii)
plot(hawaii.ts)

head(hawaii)

# Time series plot
autoplot(hawaii.ts, main = "Hawaii CO2", xlab = "Time", ylab = "Number")

plot.ts(hawaii.ts, main = "Hawaii CO2", xlab = "Time", ylab = "Number")
abline(reg=lm(hawaii.ts~time(hawaii.ts)), col = "red")


# Identifying outliers
Out <- boxplot(hawaii.ts)$out
length(which(hawaii.ts %in% Out))

den <- density(hawaii.ts, adjust = 1)
plot(den, main = "density plot for Hawaii CO2")
polygon(den, col = "red", border = "blue")

outs <- tso(hawaii.ts, types = c("TC", "AO", "LS", "IO", "SLS"))
tsoutliers(hawaii.ts)
plot(outs)

# Another inspection
time(hawaii.ts)
cycle(hawaii.ts)

summary(hawaii.ts)
str(hawaii.ts)
dim(hawaii.ts)
sd(hawaii.ts)
```

#Analysis
```{r analysis}
#Initial check
adf.test(hawaii.ts)
#kpss.test(hawaii.ts)






#Trend OR Seasonality
decomp <- decompose(hawaii.ts, type=c("additive")) # use type = "additive" for additive components by MA
plot(decomp) 
decomp2 <- decompose(hawaii.ts, type=c("multiplicative")) # use type = "additive" for additive components by MA
plot(decomp2) 

lowesdecomp <- stl(hawaii.ts, s.window = "periodic") #seasonal decomposition by lowess
lowesdecomp

plot(seasadj(lowesdecomp))
plot(seasadj(decomp2))
plot(seasadj(decomp))

seasonplot(seasadj(lowesdecomp), col = rainbow(10), main = "seasonal plot")
seasonplot(hawaii.ts, col = rainbow(10), main = "seasonal plot for original data")


ndiffs(hawaii.ts) #check how many differencing is needed 
ndiffs(seasadj(decomp2))
ndiffs(seasadj(lowesdecomp))

plot(diff(hawaii.ts, 12, 1), type = "o")
mean(diff(hawaii.ts, 12, 1))
adf.test(diff(hawaii.ts, 12, 1))

adf <- 0 #Initializing vector for adf data 
for(i in 1:11){
  adf[i] <- adf.test(diff(diff(hawaii.ts, 12, i)))$statistic
}

which.min(adf) #check which model is the most significant/most stationary

plot(diff(diff(hawaii.ts, 12, 1)), type = "o")
mean(diff(diff(hawaii.ts, 12, 1)))







#It is a new dataset after 1st and seasonal differencings
newdata <- diff(diff(hawaii.ts, 12, 1))


#Moving average smoothings/filtering
mv <- stats::filter(newdata, sides = 2, filter = c(0.5, rep(1, 11), 0.5)/12)
plot(newdata, main = "Detrending")
lines(mv, col = 2)

#Kernel smooth
plot(newdata)
lines(ksmooth(time(newdata), newdata, kernel = "normal", bandwidth = 1), col = 3)

#Lowess 
plot(newdata)
lines(lowess(newdata), col = 4)

#smooth splines
lines(smooth.spline(time(newdata), newdata, spar = 0.5), col = 5)







#Scatterplot matrix and acf and pacf for sanity check
lag1.plot(hawaii.ts, max.lag = 12)
acf(hawaii.ts)
pacf(hawaii.ts)

lag1.plot(newdata, max.lag = 12)
```

#(S)Arima
```{r arima}
#Split the train and testing
test <- tail(hawaii.ts, 12)
train <- head(hawaii.ts, -12)

length(test)
length(train)
length(hawaii.ts)

newtrain <- diff(diff(train, 12, 1))
  




#ACF, PACF, EACF
acf2(newtrain, main = "ACF and PACF for transformed Hawaii CO2")
eacf(newtrain)






#Sarima - model comparison (out = not appropriate, good = appropriate)

sarima(train, 2, 1, 2, 0, 1, 1, 12) #out
sarima(train, 2, 1, 1, 0, 1, 1, 12) #good 
sarima(train, 1, 1, 2, 0, 1, 1, 12) #out


sarima(train, 0, 1, 2, 0, 1, 1, 12) #out
sarima(train, 1, 1, 1, 0, 1, 1, 12) #good 
sarima(train, 0, 1, 1, 0, 1, 1, 12) #good

sarima(train, 1, 1, 0, 0, 1, 1, 12) #out
sarima(train, 0, 1, 0, 0, 1, 1, 12) #out




sarima(train, 2, 1, 2, 1, 1, 1, 12) #out
sarima(train, 2, 1, 1, 1, 1, 1, 12) #out
sarima(train, 1, 1, 2, 1, 1, 1, 12) #out


sarima(train, 0, 1, 2, 1, 1, 1, 12) #out
sarima(train, 1, 1, 1, 1, 1, 1, 12) #out
sarima(train, 0, 1, 1, 1, 1, 1, 12) #out

sarima(train, 1, 1, 0, 1, 1, 1, 12) #out
sarima(train, 0, 1, 0, 1, 1, 1, 12) #out

#turning.point.test(train)
#shapiro.test()
#qqnorm()






#Model parameter estimation
model1 <- sarima(train, 2, 1, 1, 0, 1, 1, 12)
model2 <- sarima(train, 1, 1, 1, 0, 1, 1, 12) 
model3 <- sarima(train, 0, 1, 1, 0, 1, 1, 12) 

model1$ttable[,1]
model2$ttable[,1]
model3$ttable[,1]




#Forecasting
model1_for <- sarima.for(train, 12, 2, 1, 1, 0, 1, 1, 12)
lines(test, type = "o", col = "blue") #real 
title(main = "model 1 forecasting (red = predict, blue = real)")

model2_for <- sarima.for(train, 12, 1, 1, 1, 0, 1, 1, 12)
lines(test, type = "o", col = "blue") #real 
title("model 2 forecasting (red = predict, blue = real)")

model3_for <- sarima.for(train, 12, 0, 1, 1, 0, 1, 1, 12)
lines(test, type = "o", col = "blue") #real 
title("model 3 forecasting (red = predict, blue = real)")



model_for <- sarima.for(hawaii.ts, 24, 0, 1, 1, 0, 1, 1, 12)
lines(test, type = "o", col = "blue") 
title("Future 2-year forecasting")



#Prediction and Standard Error table
predict_data <- data.frame(model1_prediction = model1_for$pred,
                     model1_sd = model1_for$se,
                     model2_prediction = model2_for$pred,
                     model2_sd = model2_for$se,
                     model3_prediction = model3_for$pred,
                     model3_sd = model1_for$se)
predict_data <- round(predict_data, 3)
predict_data





#MSE and Mean MSE table
mse_data <- matrix(0, 12, 3)

for(j in 1:3){
  for(i in 1:12){
    mse_data[i, j] <- (test[i] - predict_data[i, 2*j - 1])^2
  }
}
mse_data <- rbind(mse_data, colMeans(mse_data))
colnames(mse_data) <- c("Model1", "Model2", "Model3")
name <- 0
for(i in 1:nrow(mse_data) - 1){
  name[i] <- paste0("MSE: 1990, ", i)
}
name <- c(name, "Mean MSE")
rownames(mse_data) <- name

mse_data
```

#Spectral Analysis
```{r}
#Good reference: http://www.stat.pitt.edu/stoffer/tsa4/R_toot.htm
head(newdata,20)
```


#Various Detrending Method

## detrended by differencing
```{r}
plot(newdata)   


#pick the key frequencies
final <- spec.pgram(newdata, kernel("modified.daniell", c(3, 3)),
                    taper=0.1, plot = F)

keyidx <- c(1, which(diff(sign(diff(final$spec)))==-2) + 1)
keyfreq <- final$freq[keyidx]
abline(v=keyfreq, lty=2)


topfreq <- keyfreq[order(final$spec[keyidx], decreasing = T)][1:3]/12
time <- 1: length(hawaii$Carbondioxide)

terms <- do.call(cbind, lapply(topfreq, function(freq) {
  cbind(cos(2 * pi * freq * time), sin(2 * pi * freq * time))
}))



combined <- data.frame(hawaii = hawaii$Carbondioxide, time, terms)
fit <- lm(hawaii ~ ., combined)
summary(fit)
plot(time, hawaii$Carbondioxide, type="l")
lines(time, fit$fitted.values, lty=2, col="red")

plot(1:50, hawaii$Carbondioxide[1:50], type="l")
lines(1:50, fit$fitted.values[1:50], lty=2, col="red")


```

##Detrended by lm
```{r}

t <- 1:length(hawaii.ts)
fit <- lm(hawaii.ts ~ t)
hawaii_detrended <- fit$residuals
plot.ts(hawaii_detrended)

fit <- lm(hawaii.ts ~ t + I(t^2))
hawaii_detrended <- fit$residuals
plot.ts(hawaii_detrended)


```

##Detrended by stl
```{r}

trend = stl(hawaii.ts, s.window = "periodic")$time.series[,2]

detrended_stl <- hawaii.ts - (trend - trend[1])

#length(detrended_stl)-12

newdata = detrended_stl[1 : 372]

plot(detrended_stl)


```


```{r}
# Effect of different span on Kernel


k0.smooth <- spec.pgram(newdata, log='no', taper=0, pad=0, fast=FALSE, demean=TRUE, detrend=FALSE, plot = F) 

autoplot(k0.smooth, log = 'no', main = "Raw periodogram")


k1 <- kernel("daniell", c(1, 1))
k1.smooth <- spec.pgram(newdata, kernel = k1, log = 'no', taper = 0, plot = FALSE)
smooth.df <- data.frame(freq = k1.smooth$freq, `c(1,1)` = k1.smooth$spec)
names(smooth.df) <- c("frequency", "c(1,1)")
# Add other smooths
k2 <- kernel("daniell", c(3, 3))
smooth.df[, "c(3,3)"] <- spec.pgram(newdata, kernel = k2, log = 'no', taper = 0, plot = FALSE)$spec
k3 <- kernel("daniell", c(5, 5))
smooth.df[, "c(5,5)"] <- spec.pgram(newdata, kernel = k3, log = 'no', taper = 0, plot = FALSE)$spec

k4 <- kernel("daniell", c(7, 7))
smooth.df[, "c(7,7)"] <- spec.pgram(newdata, kernel = k4, log = 'no', taper = 0, plot = FALSE)$spec

# Melt dataframe in order to plot three graph together

smooth.df1 <- melt(smooth.df, variable.name = "dimension", 
                  id.vars = "frequency", value.name = "spectrum")
plot1 <- ggplot(data = subset(smooth.df1)) + geom_path(aes(x = frequency, y = spectrum, color = dimension)) +  scale_x_continuous("frequency")


plot2 <- ggplot(data = subset(smooth.df1)) + 
  geom_path(aes(x = frequency, y = spectrum, color = dimension)) + 
  scale_x_continuous("frequency") +  scale_y_log10() + 
  labs(title = "log spacing on spectrum")

grid.arrange(plot1, plot2)


```


***k1 looks smooth enough and has reasonable number of peaks

```{r}
# Effect of different span(dimention) on Kernel

k1 <- kernel("daniell", c(3, 3, 3))
k1.smooth <- spec.pgram(newdata, kernel = k1, log = 'no', taper = 0, plot = FALSE)
smooth.df <- data.frame(freq = k1.smooth$freq, `c(3,3,3)` = k1.smooth$spec)
names(smooth.df) <- c("frequency", "c(3,3,3)")
# Add other smooths
k2 <- kernel("daniell", c(3,3))
smooth.df[, "c(3,3)"] <- spec.pgram(newdata, kernel = k2, log = 'no', taper = 0, plot = FALSE)$spec
k3 <- kernel("daniell", c(3))
smooth.df[, "c(3)"] <- spec.pgram(newdata, kernel = k3, log = 'no', taper = 0, plot = FALSE)$spec



# Melt dataframe in order to plot three graph together

smooth.df <- melt(smooth.df, variable.name = "dimension", 
                  id.vars = "frequency", value.name = "spectrum")
plot1 <- ggplot(data = subset(smooth.df)) + geom_path(aes(x = frequency, y = spectrum, color = dimension)) +  scale_x_continuous("frequency") 
plot2 <- ggplot(data = subset(smooth.df)) + 
  geom_path(aes(x = frequency, y = spectrum, color = dimension)) + 
  scale_x_continuous("frequency") +  scale_y_log10() + 
  labs(title = "log spacing on spectrum")

grid.arrange(plot1, plot2)

```

```{r, warning=FALSE}
#Effect of different kernel

k1 <- kernel("daniell", 3, 1)
k1.smooth <- spec.pgram(newdata, kernel = k1, log = 'no', taper = 0, plot = FALSE)
smooth.df <- data.frame(freq = k1.smooth$freq, `daniell` = k1.smooth$spec)
names(smooth.df) <- c("frequency", "daniell")
# Add other smooths kernel
k2 <- kernel("dirichlet", 3, 1)  
smooth.df[, "dirichlet"] <- spec.pgram(newdata, kernel = k2, log = 'no', taper = 0, plot = FALSE)$spec
k3 <- kernel("fejer", 3, 1)
smooth.df[, "fejer"] <- spec.pgram(newdata, kernel = k3, log = 'no', taper = 0, plot = FALSE)$spec
k4 <- kernel("modified.daniell", 3, 1)
smooth.df[, "modi.daniell"] <- spec.pgram(newdata, kernel = k4, log = 'no', taper = 0, plot = FALSE)$spec


# Melt dataframe in order to plot three graph together
smooth.df <- melt(smooth.df, variable.name = "kernel", 
                  id.vars = "frequency", value.name = "spectrum")

which(is.nan(smooth.df[,3]))


plot1 <- ggplot(data = subset(smooth.df)) + geom_path(aes(x = frequency, y = spectrum, color = kernel)) +  scale_x_continuous("frequency") 

plot2 <- ggplot(data = subset(smooth.df)) + 
  geom_path(aes(x = frequency, y = spectrum, color = kernel)) + 
  scale_x_continuous("frequency") +  scale_y_log10() + 
  labs(title = "log spacing on spectrum")

grid.arrange(plot1, plot2)


```

```{r}

# Effect of different tapering - less important the longer your time series is, but it can be very important in short series


k1 <- kernel("modified.daniell", c(3,3) , 2)

k1.smooth <- spec.pgram(newdata, kernel = k1, taper = 0, log = 'no', plot = FALSE)
smooth.df <- data.frame(freq = k1.smooth$freq, `0%` = k1.smooth$spec)
names(smooth.df) <- c("frequency", "0%")
# Add other tapers
smooth.df[, "10%"] <- spec.pgram(newdata, kernel = k1, taper = 0.1, log = 'no', plot = FALSE)$spec
smooth.df[, "30%"] <- spec.pgram(newdata, kernel = k1, taper = 0.3, log = 'no',  plot = FALSE)$spec

smooth.df <- melt(smooth.df, variable.name = "taper", 
                  id.vars = "frequency", value.name = "spectrum")
plot1 <- ggplot(data = subset(smooth.df)) + geom_path(aes(x = frequency, y = spectrum, color = taper)) + scale_x_continuous("frequency") 
# +  scale_y_log10()

plot2 <- ggplot(data = subset(smooth.df)) + geom_path(aes(x = frequency, y = spectrum, color = taper)) + scale_x_continuous("frequency") + scale_y_log10() + labs(title = "log spacing on spectrum")

grid.arrange(plot1, plot2)


```

## Pick the key frequencies and compare the generated features with the original dataset
```{r}

#pick the key frequencies
final <- spec.pgram(newdata, kernel("modified.daniell", c(3, 3)), taper = 0.1)
keyidx <- c(1, which(diff(sign(diff(final$spec)))==-2) + 1)
keyfreq <- final$freq[keyidx]
abline(v=keyfreq, lty=2)

# compare with the parametric spectral estimator 
plot(final)
AR <- spec.ar(newdata, plot=F)
lines(AR$freq, AR$spec, lty=2, col="red")

# pick top three frequencies and use these to generate features in terms of sin and cos functions. 


topfreq <- keyfreq[order(final$spec[keyidx], decreasing = T)][1:2]

cat("Top frequencies are :", topfreq[1], " and ", topfreq[2])

time <- 1: length(newdata)

terms <- do.call(cbind, lapply(topfreq, function(freq) {
  cbind(cos(2 * pi * freq * time), sin(2 * pi * freq * time))
}))



combined <- data.frame(hawaii = hawaii$Carbondioxide[1:length(time)], time, terms)
fit <- lm(hawaii ~ ., combined)
summary(fit)
plot(time, hawaii$Carbondioxide[1:length(time)], type="l")
lines(time, fit$fitted.values, lty=2, col="red")

plot(1:50, hawaii$Carbondioxide[1:50], type="l")
lines(1:50, fit$fitted.values[1:50], lty=2, col="red")


```

## Forecast 12 observations
```{r}

time_new <- (length(time) + 1):(length(time) + 12)

terms_new <- do.call(cbind, lapply(topfreq, function(freq) {
  cbind(cos(2 * pi * freq * time_new), sin(2 * pi * freq * time_new))
}))

combined_new <- data.frame(time_new, terms_new)
colnames(combined_new) <- colnames(combined)[-1]
predictions <- predict.lm(fit, newdata=combined_new,interval="prediction", level=.95)


plot(time, hawaii$Carbondioxide[1:length(time)], type="l", xlim=c(0, 390), ylim=c(310, 360))
lines(time, fit$fitted.values, lty=2, col="red")
lines(time_new, predictions[, "fit"], col="blue")
matlines(time_new, predictions[, 2:3], col = "purple", lty=3)


#Forecasted data Enlarged
plot(350:372, hawaii$Carbondioxide[350:length(time)], type="l", xlim=c(350, 386), ylim=c(342, 358))
lines(350:372, fit$fitted.values[350:372], lty=2, col="red")
lines(time_new, predictions[, "fit"], col="blue")
matlines(time_new, predictions[, 2:3], col = "purple", lty=3)

```

## Calculating MSE from spectral model
```{r}

MSE_spectral <- mean((hawaii$Carbondioxide[373:384] - predictions[, "fit"])^2)

MSE_spectral

```

## Predicting future 24 time points from the final spectral model
```{r}

time <- 1: length(hawaii$Carbondioxide)

terms <- do.call(cbind, lapply(topfreq, function(freq) {
  cbind(cos(2 * pi * freq * time), sin(2 * pi * freq * time))
}))

combined <- data.frame(hawaii = hawaii$Carbondioxide, time, terms)
fit <- lm(hawaii ~ ., combined)

time_new <- 385:408

terms_new <- do.call(cbind, lapply(topfreq, function(freq) {
  cbind(cos(2 * pi * freq * time_new), sin(2 * pi * freq * time_new))
}))

combined_new <- data.frame(time_new, terms_new)
colnames(combined_new) <- colnames(combined)[-1]

predictions <- predict.lm(fit, newdata=combined_new, interval="prediction", level=.95)


plot(time, hawaii$Carbondioxide[1:length(time)], type="l", xlim=c(0, 410), ylim=c(310, 360))
lines(time, fit$fitted.values, lty=2, col="red")
lines(time_new, predictions[, "fit"], col="blue")
matlines(time_new, predictions[, 2:3], col = "purple", lty=3)
```





